\section{Related Work}
\label{sec:related}

\textbf{Runtime oversight of agents.}
A growing line of work interrupts an agent by checking its trajectory against
safety conditions. AgentSpec~\cite{agentspec} compiles user-specified rules
(trigger--predicate--enforce) and fires when a predicate matches the current
action; AgentHarm~\cite{agentharm} benchmarks harmful agent behavior that such
monitors aim to catch. These systems threshold per-action predicates rather
than an accumulated, time-decayed statistic, and so lie outside the
leaky-integrator class our result concerns. ProbGuard~\cite{probguard} is the
closest in spirit---it predicts the probability of reaching an unsafe state and
intervenes when that probability crosses a threshold---but the quantity it
thresholds is an \emph{instantaneous} predicted probability recomputed from the
current symbolic state at each step, not a statistic that integrates past
evidence with a wall-clock decay. It is therefore also outside our class: it
has no half-life to mis-calibrate against cadence, and our bistability result
does not apply to it. Our prior work~\cite{P1} is the wall-clock-calibrated
member whose failure this paper dissects.

\textbf{Sequential change detection.}
The classical home of edge-versus-level detection is the cumulative-sum chart
of Page~\cite{page1954}, with asymptotic optimality established by
Lorden~\cite{lorden1971} and the modern landscape surveyed by Truong et
al.~\cite{truong2020}. CUSUM is defined in \emph{sample time}---its drift term
is subtracted once per observation---which is exactly why our Section~7 control
(I2) is cadence-invariant: it is classical CUSUM behaving classically. We claim
no novelty for edge detection or for CUSUM; the contribution is the
\emph{calibration-mismatch} analysis that separates sample-time from
wall-clock-time dynamics at \emph{agent} cadence, where inter-action intervals
vary across deployments by orders of magnitude. The two-threshold hysteresis
our T3 trigger uses is the Schmitt trigger~\cite{schmitt1938}, the canonical
bistable detector with separate set/reset levels.

\textbf{Calibrated-decay monitors and label variation.}
Wall-clock-calibrated leaky integrators---exponential-moving-average behavioral
baselines and drift statistics with fixed half-lives---are common in production
agent and service monitoring; they are precisely the family our analysis places
at risk when the sampling cadence is not the cadence the half-life was tuned
for. Finally, our evaluation of whether any trigger \emph{recovers human
intervention timing} inherits the framing of LLM-as-judge
evaluation~\cite{zheng2023} and of human label
variation~\cite{plank2022,krippendorff2004}: as in~\cite{P1}, the target itself
is a low-reliability construct, so repairing the detector (Section~9) does not,
and cannot, repair the target.
