\documentclass{article}

% --- NeurIPS-workshop-style preprint ---
% NOTE: neurips_2024.sty here is a self-contained minimal stand-in (the
% official style file is not redistributable in this environment). For
% submission, drop in the official neurips_2024.sty from the CfP; the body
% below uses only standard macros and compiles against either.
\usepackage[preprint]{neurips_2024}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{natbib}
\graphicspath{{figures/}}

\hypersetup{colorlinks=true, linkcolor=blue!50!black,
            citecolor=blue!50!black, urlcolor=blue!50!black}

\title{Bistable by Construction: Wall-Clock-Calibrated State Monitors\\
Have No Moment-Detection Regime at Agent Cadence}

\author{%
  Manvendra Modgil\\
  Modint Intelligence\\
  \texttt{manvendramodgil.ai@gmail.com}\\
}

\begin{document}
\maketitle

\begin{abstract}
Runtime monitors for autonomous agents commonly threshold an accumulated
internal state---a behavioral baseline, a drift statistic, or, in our prior
work, a modeled affective state. We previously reported a State Saturation
Trap: threshold-on-state triggers over a continuous affect engine become
near-constant alarms on SWE-bench debugging agents~\cite{P1}. A post-release
audit of our replay pipeline found that the engine received $\Delta t = 0$
between actions, so its exponential decay never operated: the published trap is
a pure-accumulator result. We correct the record (erratum in \cite{P1}-v2) and
treat the flaw as an experiment design. The key variable it exposes is one the
monitoring literature does not track: whether the monitor's dynamics are
calibrated in \emph{sample time} (per observation, as in classical CUSUM) or in
\emph{wall-clock time} (half-lives in seconds, as in affect models and EMA
baselines). On fixed-rate streams these coincide; on agent streams---where
inter-action time varies by orders of magnitude across deployments---they do
not. A pre-registered sweep over uniform inter-action intervals
($\Delta t \in \{0..600\}$\,s) on 20 trajectories shows the
wall-clock-calibrated level trigger has exactly two regimes: at
$\Delta t \le 1$\,s it is a constant alarm (trap holds 20/20; median 18 firings
per run); at $\Delta t \ge 60$\,s it is silent (the state never reaches
threshold). Every trajectory's critical $\Delta t$ lies in $(1, 30]$\,s.
Hook-instrumented real agent runs measure inter-action latency at median
1.53\,s (p90 2.33\,s)---real autonomous coding cadence sits firmly inside the
trap regime, vindicating \cite{P1}'s empirical finding under a corrected
mechanism---and replaying real ordered latency sequences shows the regimes are
stable within a run (19/20 clean single-crossing accumulators; bounded flicker
only when bursts are scaled into the critical band, exploratory). The regime is
therefore selected by the deployment's latency profile, not by within-run
noise. The structure is a property of the calibration class, not the engine: a
minimal wall-clock leaky accumulator over the raw per-action error stream
reproduces the same cliff (20/20 trap at $\Delta t = 0$; 0/20 crossings at
$\Delta t \ge 60$; critical band of the same order of magnitude, Spearman
$\rho = 0.530$ with the engine's), while a sample-time CUSUM over the identical
stream is exactly invariant across the entire $\Delta t$ grid (20/20 identical
fire patterns). A rising-edge trigger with hysteresis at the identical
threshold fires 0--3 times per trajectory in every condition tested, including
a live instrumented demo. A zero-parameter balance model,
$r \ge (1 - 2^{-\Delta t/T_{1/2}})(\theta - B)$, captures the transition's order
of magnitude but not per-trajectory variation. We conclude that
wall-clock-calibrated leaky-integrator monitors admit no operating regime in
which they act as moment detectors on agent streams; transition detection
escapes the trap at every cadence, though---consistent with \cite{P1}'s
inter-rater results---it does not thereby recover human intervention timing.
\end{abstract}

\input{sections/body}

\bibliographystyle{plainnat}
\bibliography{references}

\appendix
\input{sections/appendix}

\end{document}
